---
title: "model performance water table depth recent prediction"
author: "Alokesh Manna"
date: '2023-08-13'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##The purpose of making it is that in the previous predictions we just used a middle chunk. Used least square.
##That was not a model probably we will be finally looking for. We shall use our recent prediction with an elastic net model.

---
title: "model performance water table depth"
author: "Alokesh Manna"
date: '2023-07-27'
output: word_document
---



```{r setup, include=FALSE,fig.width=1.6,fig.height=1.6}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
```
##Big var model<https://github.com/wbnicholson/BigVAR>, <https://cran.r-project.org/web/packages/BigVAR/vignettes/BigVAR.html#install>

```{r}
#install.packages("devtools")
library(devtools)

#install_github("wbnicholson/BigVAR/BigVAR")
 #install.packages("BigVAR")
library("BigVAR")
library(palmerpenguins)
#library(tidyverse)
library(ggplot2)
library(dplyr)
#install.packages("ggpmisc")
#library(ggpmisc)
#install.packages("hydroGOF")
library("hydroGOF")
```

##Here we save all the data that we have for 3 weather station.
```{r}
# save(dat_ws80_2006_2019,file="/Users/alokesh.manna17gmail.com/NSF_internship_2023/raw_data/dat_ws80_2006_2019.Rda")
# save(dat_ws77_2006_2019,file="/Users/alokesh.manna17gmail.com/NSF_internship_2023/raw_data/dat_ws77_2006_2019.Rda")
 #save(dat_ws78_2006_2019,file="/Users/alokesh.manna17gmail.com/NSF_internship_2023/raw_data/dat_ws78_2006_2019.Rda")
```
##Load the data and see the columns

```{r}
load("~/NSF_internship_2023/raw_data/dat_ws77_2006_2019.Rda")
load("~/NSF_internship_2023/raw_data/dat_ws78_2006_2019.Rda")
load("~/NSF_internship_2023/raw_data/dat_ws80_2006_2019.Rda")
colnames(dat_ws77_2006_2019)
colnames(dat_ws78_2006_2019)
colnames(dat_ws80_2006_2019)

##Data preparation for bigvar. First 2 columns are 2 outputs of Y variable. Well depth and depth measure.
library(dplyr)
Y_77=dat_ws77_2006_2019%>%dplyr::select(c(Depth_cm_b,elev_m_asl,Air_Temp_C,Soil_Temp_,Rainfall,DailyFlow,Solar_rad_,net_rad,PET,RH,Windspeed_,Wind_direc,Vapor_Kpa))

Y_78=dat_ws78_2006_2019%>%dplyr::select(c(Depth_cm_b,Elev_m_asl,Air_Temp_C,Soil_Temp,Rainfall_m.x,Daily_Flow_newdata,Solar_rad_,net_rad,PET,RH,Windspeed_,Wind_direc,Vapor_Kpa))

Y_80=dat_ws80_2006_2019%>%dplyr::select(c(Water.Table.Depth.BGS..cm.,Elevation.of.water.table.ASL..m.,Ave_air_T,Soil_Temp_,Rainfall_m,DailyFlow_,Ave_solar,net_rad,PET,RH,Windspeed_,Wind_direc,Ave_vapor_))
```
```{r}
dim(Y_77)
dim(Y_78)
dim(Y_80)
```
```{r}
year_array=c(2006:2019)
month_array=c(1:12)
```

```{r}
Y_80_monthly=dat_ws80_2006_2019%>%dplyr::select(c(Water.Table.Depth.BGS..cm.,Elevation.of.water.table.ASL..m.,Ave_air_T,Soil_Temp_,Rainfall_m,DailyFlow_,Ave_solar,net_rad,PET,RH,Windspeed_,Wind_direc,Ave_vapor_,month_extracted,year_extracted))%>%drop_na()

monthly_data_ws80_monthlystack=array(rep(0,12*16*14),dim=c(12,16,14))##monthly_data_ws77_monthlystack[,,j] is the matrix for the month month_array[j],last and first 12 months 14 years
colnames(monthly_data_ws80_monthlystack)=c(colnames(Y_80_monthly),"rain")
for(i in 1:14){
  for(j in 1:12){
    dat_filtered=Y_80_monthly%>%filter(year_extracted==year_array[i])%>%filter(month_extracted==j)
    dat_temp=dat_filtered%>%group_by(month_extracted)%>%colMeans()
    #data.frame(dat_temp)%>%colMeans()
    rain=as.numeric(dat_temp["Rainfall_m"])*dim(dat_filtered)[1]##Rain cannot ve averaged, so again sum
 monthly_data_ws80_monthlystack[j,,i]= c(as.numeric(dat_temp),rain)
 #print(i)
# print(j)
  }
}
#head(monthly_data_ws77_monthlystack)
#monthly_data_ws80_monthlystack[,,5]##should print May data for all years ws 77
#Now we stack all the data monthly
B=NULL
for (i in 1:14){ 
  A<-rbind(B,monthly_data_ws80_monthlystack[,,i])
  B=A
}
head(B)
dim(B)
```











##Here we just take the depth not the water table height.

```{r}
COEF_77=NULL
COEF_78=NULL
COEF_80=NULL

```

```{r}
preformance_test_depth_77=NULL
preformance_test_depth_78=NULL
preformance_test_depth_80=NULL
```

```{r}
preformance_train_depth_77=NULL
preformance_train_depth_78=NULL
preformance_train_depth_80=NULL
```


```{r}
#Y_77=dat_ws77_2006_2019%>%dplyr::select(c(elev_m_asl,Air_Temp_C,Soil_Temp_,Rainfall,Avg_FlowRa,Solar_rad_,net_rad,PET,RH,Windspeed_,Wind_direc,Vapor_Kpa))
Y_77=dat_ws77_2006_2019%>%drop_na%>%dplyr::select(c(Depth_cm_b,Air_Temp_C,Soil_Temp_,Rainfall,DailyFlow,Solar_rad_,net_rad,PET,RH,Windspeed_,Wind_direc,Vapor_Kpa))
VARX=list()
VARX$k=1 # indicates that the first two series are modeled
VARX$m=11
VARX$s=2 # sets 2 as the maximal lag order for exogenous series
Model1_77=constructModel(data.matrix(Y_77),p=4,struct='BasicEN',gran=c(500,10),verbose=FALSE,VARX=VARX,T1=floor(nrow(Y_77)/3),T2=floor(2*nrow(Y_77)/3))

results=cv.BigVAR(Model1_77)
#results
#plot(results)
#SparsityPlot.BigVAR.results(results)
#predict(results,n.ahead=1, confint=TRUE)
#coef(results)
COEF_77=COEF_77%>%bind_rows(coef(results))
```

```{r}
dim(Y_77)
length(results@fitted)##You must get p differences as the traning starts p step ahead from begining. 
length(results@preds)
length(results@resids)
sum((results@resids))
#length(results@preds)+length(results@resids)
#dim(as.matrix(Y_77%>%drop_na()))
#sum(resids)
#c(results@fitted[1:T1],results@fitted[(T2+1):nrow(Y)])
results@preds[1360]
results@fitted[4076] ##This two values are same confirming the fact that in fitted we have the prediction values in last length(results@preds)=1360 values.
```

##Already NA is dropped.

```{r,,fig.width=2,fig.height=2}
Y=Y_77%>%dplyr::select("Depth_cm_b")%>%as.matrix()
X=Y_77%>%dplyr::select(-c("Depth_cm_b"))%>%as.matrix()
#T1=floor(nrow(Y)/3)
#T2=floor(2*nrow(Y)/3)
#X <- matrix(0,nrow=nrow(Y),ncol=ncol(Y))
#BICMSFE <- VARXForecastEval(Y,X,p=4,s=2,T1,T2,"BIC",1)##****This part gives the error.
#length(BICMSFE$pred)
#y=tail(Y,length(BICMSFE$pred))


x=results@preds
y=tail(Y,length(x)) #actual
mon=dat_ws77_2006_2019%>%drop_na%>%dplyr::select("month_extracted")
mon=tail(mon$month_extracted,length(x))

#c(results@fitted[1:T1],results@fitted[(T2+1):nrow(Y)]) 

preformance_test_depth_77=preformance_test_depth_77%>%bind_rows("daily"=data.frame(gof(y, x)))

pic=data.frame(y,x,mon)%>%mutate(Month=as.factor(mon))
ggplot(pic,aes(x,y))+
  geom_point(aes(colour=Month)) +
  scale_fill_manual(values =c(pic$Month))+
  geom_abline()+
  coord_fixed() +
  xlab("predicted")+
  ylab("act")+
  ylim(-300,100)+
  xlim(-300,100)+
  #ggtitle('water table depth y=x, in cm, daily ws77 recent_time')+
theme(plot.title = element_text(size = 10, face = "bold"),legend.spacing.y = unit(.001, 'cm'),legend.spacing.x = unit(.001, 'cm'),legend.position='bottom',legend.key.size =  unit(.03, 'cm'))

ggsave(file="water table depth y=x, in cm, daily ws77_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```

```{r}
# plot.new()
# date=tail(dat_ws77_2006_2019%>%drop_na%>%dplyr::select(c("Date_")),length(x))$Date_
# length(x)
# length(y)
# length(date)
# plot(date,x,ty="l", col="blue", pch="o", ylab="water table depth",xlab="date",ylim=c(-300,0),lty=1)
# lines(c(date),y,ty="l", col="red", pch="o",lty=2)
# legend("bottomright",                          
#        c("predicted", "act"),
#        lty=c(1,2),
#        col = c("blue","red"))
#date=dat_ws77_2006_2019%>%drop_na%>%dplyr::select(c("Date_"))
date=tail(dat_ws77_2006_2019%>%drop_na%>%dplyr::select(c("Date_")),length(x))$Date_
dat=data.frame(date,predicted= x, actual= y[,1])
ggplot(dat,aes(date))+
  #coord_fixed() +
  geom_line(aes(y=predicted, colour = "predicted")) + 
  geom_line(aes(y=actual, colour = "actual"))+
  xlab("Date")+
  ylab("water table elevation")+
  ylim(-300,0)+
  #xlim(-300,100)+
  #labs(color = Legend)+
  ggtitle('water table depth time series prediction recent time for daily ws 77')+
theme(plot.title = element_text(size = 10, face = "bold"),legend.spacing.y = unit(0, 'cm'),legend.spacing.x = unit(0, 'cm'),legend.position='right',legend.key.size =  unit(.15, 'cm'))
ggsave(file="depth in cm prediction ts plot recent time, daily ws77", device = "png",path="pictures_new_well_depth_recent_time")

```

```{r}
dat77=dat
```

```{r}

newdat=dat77

sd=sd(dat77$predicted-dat77$actual)
  
newdat$lower=dat77$predicted-3*sd
newdat$upper=dat77$predicted+3*sd

date_df=data.frame(date=dat_ws77_2006_2019%>%
                     dplyr::select(c("Date_"))|>
                     filter(Date_>=newdat$date[1])|>
                     pull(Date_)|>
                     c())

ggplot(date_df|>left_join(newdat),aes(date, predicted)) +
  geom_point(aes(y=predicted, colour = "predicted"),pch=2) + 
  geom_point(aes(y=actual, colour = "actual"),pch=4)+
  xlab("Date")+
  ylab("water table depth")+
  ylim(-250,100)+
  #xlim(-1,3)+
  #labs(color = Legend)+
  ggtitle('water table depth time series prediction recent time for daily ws 77')+
  #geom_smooth(stat = 'summary', fun.data = mean_cl_quantile(x))+
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = "Confidence Interval"), alpha = 0.4)+
  theme_minimal()+
theme(axis.title = element_text(size = 25), 
      plot.title = element_text(size = 15, face = "plain"),
      legend.spacing.y = unit(0, 'cm'),
      legend.spacing.x = unit(0, 'cm'),
      legend.position=c(0.5,0.1),
      legend.key.size =  unit(.15, 'cm'),
      axis.line = element_line(color='black'),
      plot.background = element_blank(),
      panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
    )
  
  #stat_summary(geom = "line", fun = mean) +
  #stat_summary(geom = "ribbon", fun.data = mean_cl_quantile, alpha = 0.3)

```
```{r}
gof(dat$predicted,dat$actual)
```

```{r}
mod_77_daily=lm(dat$actual~dat$predicted)
```

```{r}
summary(mod_77_daily)
```

```{r}
#cbind(x,y,BICMSFE$MSFE,x-y)
```

##Here I want to produce graphs which gives as a fascet for individual months
```{r}
pic =bind_cols(y,x,mon)%>%mutate(Month=as.factor(mon))%>%ggplot(aes(x = x, y = y,color=Month)) +
  scale_fill_manual(values = c(mon))+
  geom_point() +
  geom_abline()+
  coord_fixed() +
  xlab("predicted")+
  ylab("act")+
  xlim(-300,100)+
  ylim(-300,100)+
  #ggtitle("water table depth, cm, monthly-facet-daily ws77")+
  theme(plot.title = element_text(size = 8, face = "bold"))
 pic+ facet_wrap(~Month)
 # facet_grid(mon)
ggsave(file="water table depth, cm, monthly-facet-daily ws77_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time") 
```






```{r,fig.width=1.6,fig.height=1.6}

ggplot( data = data.frame(x,y),aes(x,y)) +
  # stat_poly_line()+
  # stat_poly_eq(use_label(c("eq","adj.R2", "f", "p")),size = 1.6) +
  xlab("predicted")+
  ylab("act")+
  xlim(-300,100)+
  ylim(-300,100)+
  #ggtitle("water table depth, in cm, daily ws77")+
  theme(plot.title = element_text(size = 8, face = "bold"))+
  geom_point()
ggsave(file="water table depth, in cm, daily ws77_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```

```{r}

library(ggplot2)

#create residual plot with title and axis labels
ggplot(data = data.frame(x,y), aes(x = x, y = y-x)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE)+
  labs( x='Fitted Values', y='Residuals')+
  theme(plot.title = element_text(size = 14, face = "bold"))
ggsave(file="Residual analysis, water table depth, in cm, daily ws77_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```

```{r,fig.width=1.6,fig.height=1.6}
#Y_78=dat_ws78_2006_2019%>%dplyr::select(c(Depth_cm_b,Elev_m_asl,Air_Temp_C,Soil_Temp,Rainfall_m.x,Daily_Flow_newdata,Solar_rad_,net_rad,PET,RH,Windspeed_,Wind_direc,Vapor_Kpa))
Y_78=dat_ws78_2006_2019%>%drop_na()%>%dplyr::select(c(Depth_cm_b,Air_Temp_C,Soil_Temp,Rainfall_m.x,Daily_Flow_newdata,Solar_rad_,net_rad,PET,RH,Windspeed_,Wind_direc,Vapor_Kpa))

VARX=list()
VARX$k=1 # indicates that the first two series are modeled
VARX$m=11
VARX$s=2 # sets 2 as the maximal lag order for exogenous series
Model1_78=constructModel(data.matrix(Y_78%>%drop_na),p=4,struct='BasicEN',gran=c(500,10),verbose=FALSE,VARX=VARX,T1=floor(nrow(Y_78)/3),T2=floor(2*nrow(Y_78)/3))

results=cv.BigVAR(Model1_78)
#results
#plot(results)
#SparsityPlot.BigVAR.results(results)
#predict(results,n.ahead=1, confint=TRUE)
#coef(results)
COEF_78=COEF_78%>%bind_rows(coef(results))
```

```{r,fig.width=2,fig.height=2}
Y=Y_78%>%drop_na()%>%dplyr::select("Depth_cm_b")%>%as.matrix()
X=Y_78%>%drop_na()%>%dplyr::select(-c("Depth_cm_b"))%>%as.matrix()
#T1=floor(nrow(Y)/3)
#T2=floor(2*nrow(Y)/3)
#X <- matrix(0,nrow=nrow(Y),ncol=ncol(Y))
#BICMSFE <- VARXForecastEval(Y,X,p=4,s=2,T1,T2,"BIC",1)##****This part gives the error.
#length(BICMSFE$pred)
#y=tail(Y,length(BICMSFE$pred))
#mon=dat_ws78_2006_2019%>%dplyr::select("month_extracted")
#mon=mon$month_extracted[((T1+1):T2)]
#y=c(Y[c(T1+1):T2,1])
#x=c(BICMSFE$pred)
x=results@preds
y=tail(Y,length(x))
mon=dat_ws77_2006_2019%>%drop_na%>%dplyr::select("month_extracted")
mon=tail(mon$month_extracted,length(x))

preformance_test_depth_78=preformance_test_depth_78%>%bind_cols(data.frame(gof(y, x)))

# plot(x,y,ylim=c(-300,100),xlim=c(-300,100),xlab="predicted",ylab="act",col=mon,pch=19)
# title('water table depth y=x, in cm, daily ws78',cex.main=.3)
# abline(0,1)
# legend("bottomright", legend = paste("Month", 1:12), col = 1:12, pch = 19, bty = "n",cex=0.3)


pic=data.frame(y,x,mon)%>%mutate(Month=as.factor(mon))
ggplot(pic,aes(x,y))+
  geom_point(aes(colour=Month)) +
  scale_fill_manual(values =c(pic$Month))+
  geom_abline()+
  coord_fixed() +
  xlab("predicted")+
  ylab("act")+
  ylim(-300,100)+
  xlim(-300,100)+
  #ggtitle('water table depth y=x, in cm, daily ws78')+
theme(plot.title = element_text(size = 10, face = "bold"),legend.spacing.y = unit(.001, 'cm'),legend.spacing.x = unit(.001, 'cm'),legend.position='bottom',legend.key.size =  unit(.03, 'cm'))
ggsave(file="water table depth y=x, in cm, daily ws78_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```
```{r}
# plot.new()
# date=tail(dat_ws78_2006_2019%>%drop_na%>%dplyr::select(c("Date_")),length(x))$Date_
# length(x)
# length(y)
# length(date)
# plot(date,x,ty="l", col="blue", pch="o", ylab="water table depth",xlab="date",ylim=c(-300,0),lty=1)
# lines(c(date),y,ty="l", col="red", pch="o",lty=2)
# legend("bottomright",                          
#        c("predicted", "act"),
#        lty=c(1,2),
#        col = c("blue","red"))
date=tail(dat_ws78_2006_2019%>%drop_na%>%dplyr::select(c("Date_")),length(x))$Date_
dat=data.frame(date,predicted= x, actual= y[,1])

ggplot(data = data.frame(date,predicted= x, actual= y[,1]),aes(date))+
  coord_fixed() +
  geom_line(aes(y=predicted, colour = "predicted")) + 
  geom_line(aes(y=actual, colour = "act"))+
  xlab("Date")+
  ylab("water table elevation")+
  ylim(-300,0)+
  #xlim(-300,100)+
  #labs(color = Legend)+
  ggtitle('water table depth time series prediction recent time for daily ws 78')+
theme(plot.title = element_text(size = 10, face = "bold"),legend.spacing.y = unit(0, 'cm'),legend.spacing.x = unit(0, 'cm'),legend.position='right',legend.key.size =  unit(.15, 'cm'))
ggsave(file="depth in cm prediction ts plot recent time, daily ws78", device = "png",path="pictures_new_well_depth_recent_time")
```
```{r}
dat78=dat
```


```{r}

newdat=dat78

sd=sd(dat78$predicted-dat78$actual)
  
newdat$lower=dat78$predicted-3*sd
newdat$upper=dat78$predicted+3*sd

date_df=data.frame(date=dat_ws78_2006_2019%>%
                     dplyr::select(c("Date_"))|>
                     filter(Date_>=newdat$date[1])|>
                     pull(Date_)|>
                     c())

ggplot(date_df|>left_join(newdat),aes(date, predicted)) +
#ggplot(newdat,aes(date, predicted)) +
  geom_point(aes(y=predicted, colour = "predicted"),pch=2) + 
  geom_point(aes(y=actual, colour = "actual"),pch=4)+
  xlab("Date")+
  ylab("water table depth")+
  ylim(-250,100)+
  #xlim(-1,3)+
  #labs(color = Legend)+
  ggtitle('water table depth time series prediction recent time for daily ws 78')+
  #geom_smooth(stat = 'summary', fun.data = mean_cl_quantile(x))+
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = "Confidence Interval"), alpha = 0.4)+
  theme_minimal()+
theme(axis.title = element_text(size = 25), 
      plot.title = element_text(size = 18, face = "plain"),
      legend.spacing.y = unit(0, 'cm'),
      legend.spacing.x = unit(0, 'cm'),
      legend.position=c(0.5,0.1),
      legend.key.size =  unit(.15, 'cm'),
      axis.line = element_line(color='black'),
      plot.background = element_blank(),
      panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
    )
  
  #stat_summary(geom = "line", fun = mean) +
  #stat_summary(geom = "ribbon", fun.data = mean_cl_quantile, alpha = 0.3)

```
A wet year plot 2016
```{r}

newdat=dat78|>filter(date<"2017-01-01")|>filter(date>"2015-12-31")

sd=sd(dat78$predicted-dat78$actual)
  
newdat$lower=newdat$predicted-3*sd
newdat$upper=newdat$predicted+3*sd

date_df=data.frame(date=dat_ws78_2006_2019%>%
                     dplyr::select(c("Date_"))|>
                     filter(Date_>=newdat$date[1])|>
                     pull(Date_)|>
                     c())

ggplot(date_df|>
         left_join(newdat)|>
         filter(date<"2017-01-01")|>
         filter(date>"2015-12-31"),aes(date, predicted)) +
#ggplot(newdat,aes(date, predicted)) +
  geom_point(aes(y=predicted, colour = "predicted"),pch=2) + 
  geom_point(aes(y=actual, colour = "actual"),pch=4)+
  xlab("Date")+
  ylab("water table depth")+
  ylim(-250,100)+
  #xlim(-1,3)+
  #labs(color = Legend)+
  ggtitle('water table depth time series prediction wet season for daily ws 78')+
  #geom_smooth(stat = 'summary', fun.data = mean_cl_quantile(x))+
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = "Confidence Interval"), alpha = 0.4)+
  theme_minimal()+
theme(axis.title = element_text(size = 25), 
      plot.title = element_text(size = 18, face = "plain"),
      legend.spacing.y = unit(0, 'cm'),
      legend.spacing.x = unit(0, 'cm'),
      legend.position=c(0.5,0.1),
      legend.key.size =  unit(.15, 'cm'),
      axis.line = element_line(color='black'),
      plot.background = element_blank(),
      panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
    )
  
  #stat_summary(geom = "line", fun = mean) +
  #stat_summary(geom = "ribbon", fun.data = mean_cl_quantile, alpha = 0.3)

```









A dry year plot 2007
```{r}

newdat=dat78|>filter(date<"2008-01-01")|>filter(date>"2006-12-31")

sd=sd(dat78$predicted-dat78$actual)
  
newdat$lower=newdat$predicted-3*sd
newdat$upper=newdat$predicted+3*sd

date_df=data.frame(date=dat_ws78_2006_2019%>%
                     dplyr::select(c("Date_"))|>
                     filter(Date_>=newdat$date[1])|>
                     pull(Date_)|>
                     c())

ggplot(date_df|>
         left_join(newdat)|>
         filter(date<"2008-01-01")|>
         filter(date>"2006-12-31"),aes(date, predicted)) +
#ggplot(newdat,aes(date, predicted)) +
  geom_point(aes(y=predicted, colour = "predicted"),pch=2) + 
  geom_point(aes(y=actual, colour = "actual"),pch=4)+
  xlab("Date")+
  ylab("water table depth")+
  ylim(-250,100)+
  #xlim(-1,3)+
  #labs(color = Legend)+
  ggtitle('water table depth time series prediction recent time for daily ws 78')+
  #geom_smooth(stat = 'summary', fun.data = mean_cl_quantile(x))+
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = "Confidence Interval"), alpha = 0.4)+
  theme_minimal()+
theme(axis.title = element_text(size = 25), 
      plot.title = element_text(size = 18, face = "plain"),
      legend.spacing.y = unit(0, 'cm'),
      legend.spacing.x = unit(0, 'cm'),
      legend.position=c(0.5,0.1),
      legend.key.size =  unit(.15, 'cm'),
      axis.line = element_line(color='black'),
      plot.background = element_blank(),
      panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
    )
  
  #stat_summary(geom = "line", fun = mean) +
  #stat_summary(geom = "ribbon", fun.data = mean_cl_quantile, alpha = 0.3)

```


```{r}
gof(dat$predicted,dat$actual)
```

```{r}
mod_78_daily=lm(dat$actual~dat$predicted)
```

```{r}
summary(mod_78_daily)
```

```{r}
pic =bind_cols(y,x,mon)%>%mutate(Month=as.factor(mon))%>%ggplot(aes(x = x, y = y,color=Month)) +
  scale_fill_manual(values = c(mon))+
  geom_point() +
  geom_abline()+
  coord_fixed() +
  xlab("predicted")+
  ylab("act")+
  xlim(-300,100)+
  ylim(-300,100)+
  #ggtitle("water table depth, cm, monthly-facet-daily ws78")+
  theme(plot.title = element_text(size = 8, face = "bold"))
 pic+ facet_wrap(~Month)
 # facet_grid(mon)
ggsave(file="water table depth, cm, monthly-facet-daily ws78_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time") 
```

```{r,fig.width=1.6,fig.height=1.6}

ggplot( data = data.frame(x,y),aes(x,y)) +
  # stat_poly_line()+
  # stat_poly_eq(use_label(c("eq","adj.R2", "f", "p")),size = 1.6) +
  xlab("predicted")+
  ylab("act")+
  xlim(-300,100)+
  ylim(-300,100)+
  #ggtitle("water table depth, in cm, daily ws78")+
  geom_point()+
  theme(plot.title = element_text(size = 8, face = "bold"))
ggsave(file="water table depth, in cm, daily ws78_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```

```{r}

library(ggplot2)

#create residual plot with title and axis labels
ggplot(data = data.frame(x,y), aes(x = x, y = y-x)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE)+
  labs(x='Fitted Values', y='Residuals')+
  theme(plot.title = element_text(size = 14, face = "bold"))
ggsave(file="Residual analysis, water table depth, in cm, daily ws78_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```


```{r,fig.width=1.6,fig.height=1.6}
Y_80=dat_ws80_2006_2019%>%drop_na()%>%dplyr::select(c(Water.Table.Depth.BGS..cm.,Ave_air_T,Soil_Temp_,Rainfall_m,DailyFlow_,Ave_solar,net_rad,PET,RH,Windspeed_,Wind_direc,Ave_vapor_))
VARX=list()
VARX$k=1 # indicates that the first two series are modeled
VARX$m=11
VARX$s=2 # sets 2 as the maximal lag order for exogenous series
Model1_80=constructModel(data.matrix(Y_80%>%drop_na),p=4,struct='BasicEN',gran=c(500,10),verbose=FALSE,VARX=VARX,T1=floor(nrow(Y_80)/3),T2=floor(2*nrow(Y_80)/3))

results=cv.BigVAR(Model1_80)
# results
# plot(results)
# #SparsityPlot.BigVAR.results(results)
# predict(results,n.ahead=1, confint=TRUE)
# coef(results)
COEF_80=COEF_80%>%bind_rows(coef(results))
```

```{r,fig.width=2,fig.height=2}
Y=Y_80%>%drop_na()%>%dplyr::select("Water.Table.Depth.BGS..cm.")%>%as.matrix()
X=Y_80%>%drop_na()%>%dplyr::select(-c("Water.Table.Depth.BGS..cm."))%>%as.matrix()
# T1=floor(nrow(Y)/3)
# T2=floor(2*nrow(Y)/3)
# #X <- matrix(0,nrow=nrow(Y),ncol=ncol(Y))
# BICMSFE <- VARXForecastEval(Y,X,p=4,s=2,T1,T2,"BIC",1)##****This part gives the error.
# BICMSFE
# length(BICMSFE$pred)

# mon=dat_ws80_2006_2019%>%dplyr::select("month_extracted")
# mon=mon$month_extracted[((T1+1):T2)]
# y=c(Y[c(T1+1):T2,1])
# x=c(BICMSFE$pred)

x=results@preds
y=tail(Y,length(x))
mon=dat_ws80_2006_2019%>%drop_na%>%dplyr::select("month_extracted")
mon=tail(mon$month_extracted,length(x))

preformance_test_depth_80=preformance_test_depth_80%>%bind_cols(data.frame(gof(y, x)))

# plot(x,y,ylim=c(-300,100),xlim=c(-300,100),xlab="predicted",ylab="act",col=mon,pch=19)
# title('water table depth y=x, in cm, daily ws80',cex.main=.5)
# abline(0,1)
# legend("bottomright", legend = paste("Month", 1:12), col = 1:12, pch = 19, bty = "n",cex=0.3)

pic=data.frame(y,x,mon)%>%mutate(Month=as.factor(mon))
ggplot(pic,aes(x,y))+
  geom_point(aes(colour=Month)) +
  scale_fill_manual(values =c(pic$Month))+
  geom_abline()+
  coord_fixed() +
  xlab("predicted")+
  ylab("act")+
  ylim(-300,100)+
  xlim(-300,100)+
  #ggtitle('water table depth y=x, in cm, daily ws80')+
theme(plot.title = element_text(size = 5, face = "bold"),legend.spacing.y = unit(0, 'cm'),legend.spacing.x = unit(0, 'cm'),legend.position='bottom',legend.key.size =  unit(.03, 'cm'))
ggsave(file="water table depth y=x, in cm, daily ws80_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```

```{r}
# plot.new()
# date=tail(dat_ws80_2006_2019%>%drop_na%>%dplyr::select(c("Date_")),length(x))$Date_
# length(x)
# length(y)
# length(date)
# plot(date,x,ty="l", col="blue", pch="o", ylab="water table depth",xlab="date",ylim=c(-300,0),lty=1)
# lines(c(date),y,ty="l", col="red", pch="o",lty=2)
# legend("bottomright",                          
#        c("predicted", "act"),
#        lty=c(1,2),
#        col = c("blue","red"))
date=tail(dat_ws80_2006_2019%>%drop_na%>%dplyr::select(c("Date_")),length(x))$Date_
dat=data.frame(date,predicted= x, actual= y[,1])

ggplot(data = data.frame(date,predicted= x, act= y[,1]),aes(date))+
  coord_fixed() +
  geom_line(aes(y=predicted, colour = "predicted")) + 
  geom_line(aes(y=act, colour = "act"))+
  xlab("Date")+
  ylab("water table elevation")+
  ylim(-300,0)+
  #xlim(-300,100)+
  #labs(color = Legend)+
  ggtitle('water table depth time series prediction recent time for daily ws 80')+
theme(plot.title = element_text(size = 10, face = "bold"),legend.spacing.y = unit(0, 'cm'),legend.spacing.x = unit(0, 'cm'),legend.position='right',legend.key.size =  unit(.15, 'cm'))
ggsave(file="depth in cm prediction ts plot recent time, daily ws80", device = "png",path="pictures_new_well_depth_recent_time")
```


```{r}
dat80=dat
```

```{r}

newdat=dat80

sd=sd(dat80$predicted-dat80$actual)
  
newdat$lower=dat80$predicted-3*sd
newdat$upper=dat80$predicted+3*sd

date_df=data.frame(date=dat_ws80_2006_2019%>%
                     dplyr::select(c("Date_"))|>
                     filter(Date_>=newdat$date[1])|>
                     pull(Date_)|>
                     c())

ggplot(date_df|>left_join(newdat),aes(date, predicted)) +
#ggplot(newdat,aes(date, predicted)) +
  geom_point(aes(y=predicted, colour = "predicted"),pch=2) + 
  geom_point(aes(y=actual, colour = "actual"),pch=4)+
  xlab("Date")+
  ylab("water table depth")+
  ylim(-250,100)+
  #xlim(-1,3)+
  #labs(color = Legend)+
  ggtitle('water table depth time series prediction recent time for daily ws 80')+
  #geom_smooth(stat = 'summary', fun.data = mean_cl_quantile(x))+
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = "Confidence Interval"), alpha = 0.4)+
  theme_minimal()+
theme(axis.title = element_text(size = 25), 
      plot.title = element_text(size = 18, face = "plain"),
      legend.spacing.y = unit(0, 'cm'),
      legend.spacing.x = unit(0, 'cm'),
      legend.position=  c(0.1,0.8)   ,#c(0.1,0.90),
      legend.key.size =  unit(.7, 'cm'),
      axis.line = element_line(color='black'),
      plot.background = element_blank(),
      panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
    )
  
  #stat_summary(geom = "line", fun = mean) +
  #stat_summary(geom = "ribbon", fun.data = mean_cl_quantile, alpha = 0.3)

```
A wet year plot 2016 for WS 80
```{r}

newdat=dat80|>filter(date<"2017-01-01")|>filter(date>"2015-12-31")

sd=sd(dat80$predicted-dat80$actual)
  
newdat$lower=newdat$predicted-3*sd
newdat$upper=newdat$predicted+3*sd

date_df=data.frame(date=dat_ws80_2006_2019%>%
                     dplyr::select(c("Date_"))|>
                     filter(Date_>=newdat$date[1])|>
                     pull(Date_)|>
                     c())

ggplot(date_df|>
         left_join(newdat)|>
         filter(date<"2017-01-01")|>
         filter(date>"2015-12-31"),aes(date, predicted)) +
#ggplot(newdat,aes(date, predicted)) +
  geom_point(aes(y=predicted, colour = "predicted"),pch=2) + 
  geom_point(aes(y=actual, colour = "actual"),pch=4)+
  xlab("Date")+
  ylab("water table depth")+
  ylim(-250,100)+
  #xlim(-1,3)+
  #labs(color = Legend)+
  ggtitle('water table depth daily level prediction for wet season 2016 ws 80')+
  #geom_smooth(stat = 'summary', fun.data = mean_cl_quantile(x))+
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = "Confidence Interval"), alpha = 0.4)+
  theme_minimal()+
theme(axis.title = element_text(size = 25), 
      plot.title = element_text(size = 18, face = "plain"),
      legend.spacing.y = unit(0, 'cm'),
      legend.spacing.x = unit(0, 'cm'),
      legend.position=c(0.1,0.9),
      legend.key.size =  unit(.15, 'cm'),
      axis.line = element_line(color='black'),
      plot.background = element_blank(),
      panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
    )
  
  #stat_summary(geom = "line", fun = mean) +
  #stat_summary(geom = "ribbon", fun.data = mean_cl_quantile, alpha = 0.3)

```




A dry year plot 2019 for WS 80
```{r}

newdat=dat80|>filter(date<"2020-01-01")|>filter(date>"2018-12-31")

sd=sd(dat80$predicted-dat80$actual)
  
newdat$lower=newdat$predicted-3*sd
newdat$upper=newdat$predicted+3*sd

date_df=data.frame(date=dat_ws80_2006_2019%>%
                     dplyr::select(c("Date_"))|>
                     filter(Date_>=newdat$date[1])|>
                     pull(Date_)|>
                     c())

ggplot(date_df|>
         left_join(newdat)|>
         filter(date<"2020-01-01")|>
         filter(date>"2018-12-31"),aes(date, predicted)) +
#ggplot(newdat,aes(date, predicted)) +
  geom_point(aes(y=predicted, colour = "predicted"),pch=2) + 
  geom_point(aes(y=actual, colour = "actual"),pch=4)+
  xlab("Date")+
  ylab("water table depth")+
  ylim(-250,100)+
  #xlim(-1,3)+
  #labs(color = Legend)+
  ggtitle('water table depth daily level prediction for dry season 2019 ws 80')+
  #geom_smooth(stat = 'summary', fun.data = mean_cl_quantile(x))+
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = "Confidence Interval"), alpha = 0.4)+
  theme_minimal()+
theme(axis.title = element_text(size = 25), 
      plot.title = element_text(size = 18, face = "plain"),
      legend.spacing.y = unit(0, 'cm'),
      legend.spacing.x = unit(0, 'cm'),
      legend.position=c(0.1,0.9),
      legend.key.size =  unit(.15, 'cm'),
      axis.line = element_line(color='black'),
      plot.background = element_blank(),
      panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
    )
  
  #stat_summary(geom = "line", fun = mean) +
  #stat_summary(geom = "ribbon", fun.data = mean_cl_quantile, alpha = 0.3)

```

```{r}
gof(dat$predicted,dat$actual)
```


```{r}
mod_80_daily=lm(dat$actual~dat$predicted)
```

```{r}
summary(mod_80_daily)
```



```{r}
pic =bind_cols(y,x,mon)%>%mutate(Month=as.factor(mon))%>%ggplot(aes(x = x, y = y,color=Month)) +
  scale_fill_manual(values = c(mon))+
  geom_point() +
  geom_abline()+
  coord_fixed() +
  xlab("predicted")+
  ylab("act")+
  xlim(-300,100)+
  ylim(-300,100)+
  ggtitle("water table depth, cm, monthly-facet-daily ws80")+
  theme(plot.title = element_text(size = 8, face = "bold"))
 pic+ facet_wrap(~Month)
 # facet_grid(mon)
ggsave(file="water table depth, cm, monthly-facet-daily ws80_recent_time.png.png", device = "png",path="pictures_new_well_depth_recent_time") 
```




```{r,fig.width=1.6,fig.height=1.6}

ggplot( data = data.frame(x,y),aes(x,y)) +
  # stat_poly_line()+
  # stat_poly_eq(use_label(c("eq","adj.R2", "f", "p")),size = 1.6) +
  xlab("predicted")+
  ylab("act")+
  xlim(-300,100)+
  ylim(-300,100)+
 
  
  #ggtitle("water table depth, in cm, daily ws80_recent_time.png")+
  geom_point()+
  theme(plot.title = element_text(size = 8, face = "bold"))
ggsave(file="water table depth, in cm, daily ws80_recent_time.png.png", device = "png",path="pictures_new_well_depth_recent_time")
```

```{r}
library(ggplot2)

#create residual plot with title and axis labels
ggplot(data = data.frame(x,y), aes(x = x, y = y-x)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE)+
  labs(x='Fitted Values', y='Residuals')+
  theme(plot.title = element_text(size = 14, face = "bold"))
ggsave(file="Residual analysis, water table depth, in cm, daily ws80_recent_time.png.png", device = "png",path="pictures_new_well_depth_recent_time")
```


```{r,fig.width=1.6,fig.height=1.6}
#plot(Y_77$elev_m_asl,Y_77$Depth_cm_b)
```




```{r}
#plot(Y_78$Depth_cm_b,Y_78$Elev_m_asl)
```
```{r}
#plot(Y_80$Water.Table.Depth.BGS..cm.,Y_80$Water.Table.Depth.BGS..cm.)
```

```{r}
#plot(Y_77$Rainfall,Y_77$Avg_FlowRa)
```

```{r}
#plot(Y_78$Rainfall_m.x,Y_78$Daily_Flow_newdata)
```

```{r}
#plot(Y_80$Rainfall_m,Y_80$DailyFlow_)
```

```{r}
#plot(Y_80$Rainfall_m[1:5111],Y_80$DailyFlow_[3:5113])

```
##Somehow rainfall does not affect directly on the well depth. I also did not find any direct relation with rainfall and daily flow. 


##Now we shall try to see this model on monthly data.

```{r,fig.width=1.6,fig.height=1.6}
Y_80_monthly=dat_ws80_2006_2019%>%drop_na()%>%dplyr::select(c(Water.Table.Depth.BGS..cm.,Elevation.of.water.table.ASL..m.,Ave_air_T,Soil_Temp_,Rainfall_m,DailyFlow_,Ave_solar,net_rad,PET,RH,Windspeed_,Wind_direc,Ave_vapor_,month_extracted,year_extracted))

monthly_data_ws80_monthlystack=array(rep(0,12*16*14),dim=c(12,16,14))##monthly_data_ws77_monthlystack[,,j] is the matrix for the month month_array[j],last and first 12 months 14 years
colnames(monthly_data_ws80_monthlystack)=c(colnames(Y_80_monthly),"rain")
for(i in 1:14){
  for(j in 1:12){
    dat_filtered=Y_80_monthly%>%filter(year_extracted==year_array[i])%>%filter(month_extracted==j)
    dat_temp=dat_filtered%>%group_by(month_extracted)%>%colMeans(na.rm = TRUE)
    #data.frame(dat_temp)%>%colMeans()
    rain=as.numeric(dat_temp["Rainfall_m"])*dim(dat_filtered)[1]##Rain cannot ve averaged, so again sum
 monthly_data_ws80_monthlystack[j,,i]= c(as.numeric(dat_temp),rain)
 #print(i)
# print(j)
  }
}
#head(monthly_data_ws77_monthlystack)
#monthly_data_ws80_monthlystack[,,5]##should print all months for 2009+5-1 data
#Now we stack all the data monthly
B_80=NULL
for (i in 1:14){ 
  A<-rbind(B_80,monthly_data_ws80_monthlystack[,,i])
  B_80=A
}
# head(B_80)
# dim(B)

dim(B_80)
#B_80[2:168 ,"Water.Table.Depth.BGS..cm."]

B_80=cbind(B_80[1:167 ,"Water.Table.Depth.BGS..cm."],data.frame(B_80)%>%dplyr::select(-c(Water.Table.Depth.BGS..cm.))%>%slice(2:168))
colnames(B_80)
```



```{r,fig.width=1.6,fig.height=1.6}
dat=B_80
#data.matrix(Y_80%>%drop_na)

VARX=list()
VARX$k=1 # indicates that the first two series are modeled
VARX$m=11
VARX$s=2 # sets 2 as the maximal lag order for exogenous series
Model2_80=constructModel(data.matrix(dat%>%dplyr::select(-c(Elevation.of.water.table.ASL..m.,Rainfall_m,month_extracted,year_extracted))%>%drop_na),p=4,struct='BasicEN',gran=c(500,10),verbose=FALSE,VARX=VARX,T1=floor(nrow(dat)/3),T2=floor(2*nrow(dat)/3))


results=cv.BigVAR(Model2_80)
# results
# plot(results)
# #SparsityPlot.BigVAR.results(results)
# predict(results,n.ahead=1, confint=TRUE)
# coef(results)
colnames(dat%>%dplyr::select(-c(Elevation.of.water.table.ASL..m.,Rainfall_m,month_extracted,year_extracted)))
COEF_80=COEF_80%>%bind_rows(coef(results))
```
#ðšƒðŸ·
#: Integer, Start of forecast evaluation period.
#ðšƒðŸ¸
#: Interger, End of forecast evaluation period.

```{r,fig.width=1.6,fig.height=1.6}
Y=dat%>%dplyr::select("B_80[1:167, \"Water.Table.Depth.BGS..cm.\"]")
X=dat%>%dplyr::select(-c(Elevation.of.water.table.ASL..m.,Rainfall_m,month_extracted,year_extracted))
# T1=floor(nrow(Y)/3)
# T2=floor(2*nrow(Y)/3)
# #X <- matrix(0,nrow=nrow(Y),ncol=ncol(Y))
# BICMSFE <- VARXForecastEval(Y,X,p=4,s=2,T1,T2,"BIC",1)##****This part gives the error.
# BICMSFE
# y=Y[c((T1+1):T2),]
# x=c(BICMSFE$pred)

x=c(results@preds)
y=tail(Y[,1],length(x))
mon=dat_ws77_2006_2019%>%drop_na%>%dplyr::select("month_extracted")
mon=tail(mon$month_extracted,length(x))


preformance_test_depth_80=preformance_test_depth_80%>%bind_cols(data.frame(gof(y, x)))

# plot(x,y,ylim=c(-300,100),xlim=c(-300,100),xlab="predicted",ylab="act",pch=19)
# title(cex.main=.5)
# abline(0,1)
# ggsave(file="water table depth y=x, in cm, monthly ws80_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```



```{r,fig.width=1.6,fig.height=1.6}

ggplot( data = data.frame(x,y),aes(x,y)) +
  # stat_poly_line()+
  # stat_poly_eq(use_label(c("eq","adj.R2", "f", "p")),size = 1.6) +
  xlab("predicted")+
  ylab("act")+
  xlim(-300,100)+
  ylim(-300,100)+
  #ggtitle("water table depth, in cm, monthly ws80")+
  geom_point()+
  theme(plot.title = element_text(size = 8, face = "bold"))
ggsave(file="water table depth, in cm, monthly ws80_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```

```{r}

library(ggplot2)

#create residual plot with title and axis labels
ggplot(data = data.frame(x,y), aes(x = x, y = y-x)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE)+
  labs( x='Fitted Values', y='Residuals')+
  theme(plot.title = element_text(size = 14, face = "bold"))
ggsave(file="Residual analysis, water table depth, in cm, monthly ws80_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```


##For ws77

```{r,fig.width=1.6,fig.height=1.6}
# Y_80_monthly=dat_ws80_2006_2019%>%dplyr::select(c(Water.Table.Depth.BGS..cm.,Water.Table.Depth.BGS..cm.,Ave_air_T,Soil_Temp_,Rainfall_m,DailyFlow_,Ave_solar,net_rad,PET,RH,Windspeed_,Wind_direc,Ave_vapor_,month_extracted,year_extracted))

Y_77_monthly=dat_ws77_2006_2019%>%drop_na()%>%dplyr::select(c(Depth_cm_b,elev_m_asl,Air_Temp_C,Soil_Temp_,Rainfall,DailyFlow,Solar_rad_,net_rad,PET,RH,Windspeed_,Wind_direc,Vapor_Kpa,month_extracted,year_extracted))

# Y_78=dat_ws78_2006_2019%>%dplyr::select(c(Depth_cm_b,Elev_m_asl,Air_Temp_C,Soil_Temp,Rainfall_m.x,Daily_Flow_newdata,Solar_rad_,net_rad,PET,RH,Windspeed_,Wind_direc,Vapor_Kpa))


monthly_data_ws77_monthlystack=array(rep(0,12*16*14),dim=c(12,16,14))##monthly_data_ws77_monthlystack[,,j] is the matrix for the month month_array[j],last and first 12 months 14 years
colnames(monthly_data_ws77_monthlystack)=c(colnames(Y_77_monthly),"rain")
for(i in 1:14){
  for(j in 1:12){
    dat_filtered=Y_77_monthly%>%filter(year_extracted==year_array[i])%>%filter(month_extracted==j)
    dat_temp=dat_filtered%>%group_by(month_extracted)%>%colMeans(na.rm = TRUE)
    #data.frame(dat_temp)%>%colMeans()
    rain=as.numeric(dat_temp["Rainfall"])*dim(dat_filtered)[1]##Rain cannot ve averaged, so again sum
 monthly_data_ws77_monthlystack[j,,i]= c(as.numeric(dat_temp),rain)
 #print(i)
# print(j)
  }
}
#head(monthly_data_ws77_monthlystack)
#monthly_data_ws80_monthlystack[,,5]##should print all months for 2009+5-1 data
#Now we stack all the data monthly
B_77=NULL
for (i in 1:14){ 
  A<-rbind(B_77,monthly_data_ws77_monthlystack[,,i])
  B_77=A
}
# head(B_80)
# dim(B)

dim(B_77)
#B_80[2:168 ,"Water.Table.Depth.BGS..cm."]

B_77=cbind(B_77[1:167 ,"Depth_cm_b"],data.frame(B_77)%>%dplyr::select(-c(Depth_cm_b))%>%slice(2:168))
```

```{r,fig.width=1.6,fig.height=1.6}
dat=B_77
#data.matrix(Y_80%>%drop_na)

VARX=list()
VARX$k=1 # indicates that the first two series are modeled
VARX$m=11
VARX$s=2 # sets 2 as the maximal lag order for exogenous series
Model2_77=constructModel(data.matrix(dat%>%dplyr::select(-c(elev_m_asl,Rainfall,month_extracted,year_extracted))%>%drop_na),p=4,struct='BasicEN',gran=c(500,10),verbose=FALSE,VARX=VARX,T1=floor(nrow(dat)/3),T2=floor(2*nrow(dat)/3))


results=cv.BigVAR(Model2_77)
# results
# plot(results)
# #SparsityPlot.BigVAR.results(results)
# predict(results,n.ahead=1, confint=TRUE)
# coef(results)
colnames(dat%>%dplyr::select(-c(elev_m_asl,Rainfall,month_extracted,year_extracted)))
COEF_77=COEF_77%>%bind_rows(coef(results))
```


```{r,fig.width=1.6,fig.height=1.6}
Y=dat%>%drop_na()%>%dplyr::select("B_77[1:167, \"Depth_cm_b\"]")
X=dat%>%drop_na()%>%dplyr::select(-c(elev_m_asl,Rainfall,month_extracted,year_extracted))
# T1=floor(nrow(Y)/3)
# T2=floor(2*nrow(Y)/3)
# #X <- matrix(0,nrow=nrow(Y),ncol=ncol(Y))
# BICMSFE <- VARXForecastEval(Y,X,p=4,s=2,T1,T2,"BIC",1)##****This part gives the error.
# BICMSFE
# #y=Y[c((T2+2):151),]
# #mon=dat_ws80_2006_2019%>%dplyr::select("month_extracted")
# #mon=mon$month_extracted[((T1+1):T2)]
# y=c(Y[c(T1+1):T2,1])
# x=c(BICMSFE$pred)
x=c(results@preds)
y=tail(Y[,1],length(x))
mon=dat_ws77_2006_2019%>%drop_na%>%dplyr::select("month_extracted")
mon=tail(mon$month_extracted,length(x))

preformance_test_depth_77=preformance_test_depth_77%>%bind_cols(data.frame(gof(y, x)))

# plot(x,y,ylim=c(-300,100),xlim=c(-300,100),xlab="predicted",ylab="act",pch=19)
# #title('water table depth y=x, in cm, monthly ws77_recent_time',cex.main=.5)
# abline(0,1)
# ggsave(file="water table depth y=x, in cm, monthly ws77_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```


```{r,fig.width=1.6,fig.height=1.6}

ggplot( data = data.frame(x,y),aes(x,y)) +
  # stat_poly_line()+
  # stat_poly_eq(use_label(c("eq","adj.R2", "f", "p")),size = 1.6) +
  xlab("predicted")+
  ylab("act")+
  xlim(-300,100)+
  ylim(-300,100)+
  #ggtitle("water table depth, in cm, monthly ws77")+
  geom_point()+
  theme(plot.title = element_text(size = 8, face = "bold"))
ggsave(file="water table depth, in cm, monthly ws77_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```


```{r}

library(ggplot2)

#create residual plot with title and axis labels
ggplot(data = data.frame(x,y), aes(x = x, y = y-x)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE)+
  #title='Residual analysis, water table depth, in cm, monthly ws77_recent_time',
  labs( x='Fitted Values', y='Residuals')+
  theme(plot.title = element_text(size = 14, face = "bold"))
ggsave(file="Residual analysis, water table depth, in cm, monthly ws77_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```

##WS78


```{r,fig.width=1.6,fig.height=1.6}
# Y_80_monthly=dat_ws80_2006_2019%>%dplyr::select(c(Water.Table.Depth.BGS..cm.,Water.Table.Depth.BGS..cm.,Ave_air_T,Soil_Temp_,Rainfall_m,DailyFlow_,Ave_solar,net_rad,PET,RH,Windspeed_,Wind_direc,Ave_vapor_,month_extracted,year_extracted))

# Y_77_monthly=dat_ws77_2006_2019%>%dplyr::select(c(Depth_cm_b,elev_m_asl,Air_Temp_C,Soil_Temp_,Rainfall,Avg_FlowRa,Solar_rad_,net_rad,PET,RH,Windspeed_,Wind_direc,Vapor_Kpa,month_extracted,year_extracted))

 Y_78_monthly=dat_ws78_2006_2019%>%dplyr::select(c(Depth_cm_b,Elev_m_asl,Air_Temp_C,Soil_Temp,Rainfall_m.x,Daily_Flow_newdata,Solar_rad_,net_rad,PET,RH,Windspeed_,Wind_direc,Vapor_Kpa,month_extracted,year_extracted))


monthly_data_ws78_monthlystack=array(rep(0,12*16*14),dim=c(12,16,14))##monthly_data_ws77_monthlystack[,,j] is the matrix for the month month_array[j],last and first 12 months 14 years
colnames(monthly_data_ws78_monthlystack)=c(colnames(Y_78_monthly),"rain")
for(i in 1:14){
  for(j in 1:12){
    dat_filtered=Y_78_monthly%>%filter(year_extracted==year_array[i])%>%filter(month_extracted==j)
    dat_temp=dat_filtered%>%group_by(month_extracted)%>%colMeans(na.rm = TRUE)
    #data.frame(dat_temp)%>%colMeans()
    rain=as.numeric(dat_temp["Rainfall_m.x"])*dim(dat_filtered)[1]##Rain cannot ve averaged, so again sum
 monthly_data_ws78_monthlystack[j,,i]= c(as.numeric(dat_temp),rain)
 #print(i)
# print(j)
  }
}
#head(monthly_data_ws77_monthlystack)
#monthly_data_ws80_monthlystack[,,5]##should print all months for 2009+5-1 data
#Now we stack all the data monthly
B_78=NULL
for (i in 1:14){ 
  A<-rbind(B_78,monthly_data_ws78_monthlystack[,,i])
  B_78=A
}
# head(B_80)
# dim(B)

dim(B_78)
#B_80[2:168 ,"Water.Table.Depth.BGS..cm."]

B_78=cbind(B_78[1:167 ,"Depth_cm_b"],data.frame(B_78)%>%dplyr::select(-c(Depth_cm_b))%>%slice(2:168))

```


```{r,fig.width=1.6,fig.height=1.6}
dat=B_78
#data.matrix(Y_80%>%drop_na)

VARX=list()
VARX$k=1 # indicates that the first two series are modeled
VARX$m=11
VARX$s=2 # sets 2 as the maximal lag order for exogenous series
Model2_78=constructModel(data.matrix(dat%>%dplyr::select(-c(Elev_m_asl,Rainfall_m.x,month_extracted,year_extracted))%>%drop_na),p=4,struct='BasicEN',gran=c(5000,10),verbose=FALSE,VARX=VARX,T1=floor(nrow(dat)/3),T2=floor(2*nrow(dat)/3))


results=cv.BigVAR(Model2_78)
# results
# plot(results)
# #SparsityPlot.BigVAR.results(results)
# predict(results,n.ahead=1, confint=TRUE)
# coef(results)
colnames(dat%>%dplyr::select(-c(Elev_m_asl,Rainfall_m.x,month_extracted,year_extracted)))
COEF_78=COEF_78%>%bind_rows(coef(results))
```
```{r,fig.width=1.6,fig.height=1.6}
Y=dat%>%drop_na()%>%dplyr::select("B_78[1:167, \"Depth_cm_b\"]")
X=dat%>%drop_na()%>%dplyr::select(-c(Elev_m_asl,Rainfall_m.x,month_extracted,year_extracted))
# T1=floor(nrow(Y)/3)
# T2=floor(2*nrow(Y)/3)
# #X <- matrix(0,nrow=nrow(Y),ncol=ncol(Y))
# BICMSFE <- VARXForecastEval(Y,X,p=4,s=2,T1,T2,"BIC",1)##****This part gives the error.
# BICMSFE
# #y=Y[c((T2+2):133),]
# #mon=dat_ws80_2006_2019%>%dplyr::select("month_extracted")
# #mon=mon$month_extracted[((T1+1):T2)]
# y=c(Y[c(T1+1):T2,1])
# x=c(BICMSFE$pred)

x=c(results@preds)
y=tail(Y[,1],length(x))
mon=dat_ws77_2006_2019%>%drop_na%>%dplyr::select("month_extracted")
mon=tail(mon$month_extracted,length(x))

preformance_test_depth_78=preformance_test_depth_78%>%bind_cols(data.frame(gof(y, x)))

# plot(x,y,ylim=c(-300,100),xlim=c(-300,100),xlab="predicted",ylab="act",pch=19)
# #title('water table depth y=x, in cm, monthly ws78_recent_time',cex.main=.5)
# abline(0,1)
# ggsave(file="water table depth y=x, in cm, monthly ws78_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```


```{r}

ggplot( data = data.frame(x,y),aes(x,y)) +
  # stat_poly_line()+
  # stat_poly_eq(use_label(c("eq","adj.R2", "f", "p")),size = 1.6) +
  xlab("predicted")+
  ylab("act")+
  xlim(-300,100)+
  ylim(-300,100)+
 
  #ggtitle("water table depth, in cm, monthly ws78")+
  geom_point()+
  theme(plot.title = element_text(size = 14, face = "bold"))
ggsave(file="water table depth, in cm, monthly ws78_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```


```{r}
library(ggplot2)

#create residual plot with title and axis labels
ggplot(data = data.frame(x,y), aes(x = x, y = y-x)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE)+
  #title='Residual analysis, water table depth, in cm, monthly ws78_recent_time',
  labs( x='Fitted Values', y='Residuals')+
  theme(plot.title = element_text(size = 14, face = "bold"))
ggsave(file="Residual analysis, water table depth, in cm, monthly ws78_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```



```{r,fig.width=1.6,fig.height=1.6}
dat
which(is.na(dat)==T)
```

###Now we shall see the seasonal data. 2 seasons (a) Growing - April1 -to Oct 30 (b) Dormant - November 1 - March 31. It will not be used but just in case we need.

```{r}
growing=c(4:10)
dormant=c(1:3,11,12)
```



```{r}
growing_dat_77=dat_ws77_2006_2019%>%filter(month_extracted%in%growing)%>%drop_na()
dormant_dat_77=dat_ws77_2006_2019%>%filter(month_extracted%in%dormant)%>%drop_na()
head(growing_dat_77)
```

```{r}
growing_dat_78=dat_ws78_2006_2019%>%filter(month_extracted%in%growing)%>%drop_na()
dormant_dat_78=dat_ws78_2006_2019%>%filter(month_extracted%in%dormant)%>%drop_na()
head(growing_dat_78)
```


```{r,fig.width=1.6,fig.height=1.6}
growing_dat_80=dat_ws80_2006_2019%>%dplyr::select(c(Water.Table.Depth.BGS..cm.,Ave_air_T,Soil_Temp_,Rainfall_m,DailyFlow_,Ave_solar,net_rad,PET,RH,Windspeed_,Wind_direc,Ave_vapor_,month_extracted))%>%filter(month_extracted%in%growing)%>%filter(month_extracted%in%growing)%>%dplyr::select(-c(month_extracted))%>%drop_na()
dormant_dat_80=dat_ws80_2006_2019%>%dplyr::select(c(Water.Table.Depth.BGS..cm.,Ave_air_T,Soil_Temp_,Rainfall_m,DailyFlow_,Ave_solar,net_rad,PET,RH,Windspeed_,Wind_direc,Ave_vapor_,month_extracted))%>%filter(month_extracted%in%dormant)%>%dplyr::select(-c(month_extracted))%>%drop_na()
head(growing_dat_80)


VARX=list()
VARX$k=1 # indicates that the first two series are modeled
VARX$m=11
VARX$s=2 # sets 2 as the maximal lag order for exogenous series
Model1_growing_80=constructModel(data.matrix(growing_dat_80%>%drop_na),p=4,struct='BasicEN',gran=c(500,10),verbose=FALSE,VARX=VARX,T1=floor(nrow(growing_dat_80)/3),T2=floor(2*nrow(growing_dat_80)/3))
Model1_dormant_80=constructModel(data.matrix(dormant_dat_80%>%drop_na),p=4,struct='BasicEN',gran=c(500,10),verbose=FALSE,VARX=VARX,T1=floor(nrow(dormant_dat_80)/3),T2=floor(2*nrow(dormant_dat_80)/3))
results_growing=cv.BigVAR(Model1_growing_80)
results_dormant=cv.BigVAR(Model1_dormant_80)
#results
# plot(results_growing)
# plot(results_dormant)
# SparsityPlot.BigVAR.results(results_growing)
# SparsityPlot.BigVAR.results(results_dormant)
# 
# predict(results_growing,n.ahead=1, confint=TRUE)
# predict(results_dormant,n.ahead=1, confint=TRUE)
# 
# coef(results_growing)
# coef(results_dormant)

COEF_80=COEF_80%>%bind_rows(coef(results_growing))
COEF_80=COEF_80%>%bind_rows(coef(results_dormant))

```


```{r,fig.width=1.6,fig.height=1.6}
Y1=growing_dat_80%>%drop_na%>%dplyr::select("Water.Table.Depth.BGS..cm.")%>%as.matrix()
X1=growing_dat_80%>%drop_na()%>%dplyr::select(-c("Water.Table.Depth.BGS..cm."))%>%as.matrix()
# T1=floor(nrow(Y1)/3)
# T2=floor(2*nrow(Y1)/3)
# #X <- matrix(0,nrow=nrow(Y),ncol=ncol(Y))
# BICMSFE <- VARXForecastEval(Y1,X1,p=4,s=2,T1,T2,"BIC",1)##****This part gives the error.
# 
# BICMSFE
# #y=tail(Y1,797)
# #mon=dat_ws80_2006_2019%>%dplyr::select("month_extracted")
# #mon=mon$month_extracted[((T1+1):T2)]
# y=c(Y1[c(T1+1):T2,1])
# x=c(BICMSFE$pred)

x=c(results_growing@preds)
y=tail(Y1[,1],length(x))
mon=dat_ws77_2006_2019%>%drop_na%>%dplyr::select("month_extracted")
mon=tail(mon$month_extracted,length(x))
preformance_test_depth_80=preformance_test_depth_80%>%bind_cols(data.frame(gof(y, x)))

# plot(x,y,ylim=c(-300,100),xlim=c(-300,100),xlab="predicted",ylab="act",pch=19)
# #title('water table depth y=x, in cm, growing ws80_recent_time',cex.main=.5)
# abline(0,1)
# ggsave(file="water table depth y=x, in cm, growing ws80_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```

```{r,fig.width=1.6,fig.height=1.6}

ggplot( data = data.frame(x,y),aes(x,y)) +
  # stat_poly_line()+
  # stat_poly_eq(use_label(c("eq","adj.R2", "f", "p")),size = 1.6) +
  xlab("predicted")+
  ylab("act")+
  xlim(-300,100)+
  ylim(-300,100)+
 
  #ggtitle("water table depth, in cm, growing ws80")+
  geom_point()+
  theme(plot.title = element_text(size = 8, face = "bold"))
ggsave(file="water table depth, in cm, growing ws80_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```


```{r}

library(ggplot2)

#create residual plot with title and axis labels
ggplot(data = data.frame(x,y), aes(x = x, y = y-x)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE)+
  labs( x='Fitted Values', y='Residuals')+
  theme(plot.title = element_text(size = 14, face = "bold"))
ggsave(file="Residual analysis, water table depth, in cm, growing ws80_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```


```{r,fig.width=1.6,fig.height=1.6}
Y2=dormant_dat_80%>%drop_na%>%dplyr::select("Water.Table.Depth.BGS..cm.")%>%as.matrix()
X2=dormant_dat_80%>%drop_na()%>%dplyr::select(-c("Water.Table.Depth.BGS..cm."))%>%as.matrix()
# T1=floor(nrow(Y2)/3)
# T2=floor(2*nrow(Y2)/3)
# #X <- matrix(0,nrow=nrow(Y),ncol=ncol(Y))
# BICMSFE <- VARXForecastEval(Y2,X2,p=4,s=2,T1,T2,"BIC",1)##****This part gives the error.
# 
# BICMSFE
# #y=tail(Y1,602)
# #mon=dat_ws80_2006_2019%>%dplyr::select("month_extracted")
# #mon=mon$month_extracted[((T1+1):T2)]
# y=c(Y2[c(T1+1):T2,1])
# x=c(BICMSFE$pred)

x=c(results_dormant@preds)
y=tail(Y2[,1],length(x))
mon=dat_ws77_2006_2019%>%drop_na%>%dplyr::select("month_extracted")
mon=tail(mon$month_extracted,length(x))
preformance_test_depth_80=preformance_test_depth_80%>%bind_cols(data.frame(gof(y, x)))

# plot(x,y,ylim=c(-300,100),xlim=c(-300,100),xlab="predicted",ylab="act",pch=19)
# #title('water table depth y=x, in cm, dormant ws80',cex.main=.5)
# abline(0,1)
# ggsave(file="water table depth y=x, in cm, dormant ws80_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```


```{r,fig.width=1.6,fig.height=1.6}

ggplot( data = data.frame(x,y),aes(x,y)) +
  # stat_poly_line()+
  # stat_poly_eq(use_label(c("eq","adj.R2", "f", "p")),size = 1.6) +
  xlab("predicted")+
  ylab("act")+
  xlim(-300,100)+
  ylim(-300,100)+
 
  #ggtitle("water table depth, in cm, dormant ws80")+
  geom_point()+
  theme(plot.title = element_text(size = 8, face = "bold"))
ggsave(file="water table depth, in cm, dormant ws80_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```

```{r}

library(ggplot2)

#create residual plot with title and axis labels
ggplot(data = data.frame(x,y), aes(x = x, y = y-x)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE)+
  #title='Residual analysis, water table depth, in cm, dormant ws80',
  labs( x='Fitted Values', y='Residuals')+
  theme(plot.title = element_text(size = 14, face = "bold"))
ggsave(file="Residual analysis, water table depth, in cm, dormant ws80_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```

```{r,fig.width=1.6,fig.height=1.6}
growing_dat_78=dat_ws78_2006_2019%>%dplyr::select(c(Depth_cm_b,Air_Temp_C,Soil_Temp,Rainfall_m.x,Daily_Flow_newdata,Solar_rad_,net_rad,PET,RH,Windspeed_,Wind_direc,Vapor_Kpa,month_extracted))%>%filter(month_extracted%in%growing)%>%dplyr::select(-c(month_extracted))
dormant_dat_78=dat_ws78_2006_2019%>%dplyr::select(c(Depth_cm_b,Air_Temp_C,Soil_Temp,Rainfall_m.x,Daily_Flow_newdata,Solar_rad_,net_rad,PET,RH,Windspeed_,Wind_direc,Vapor_Kpa,month_extracted))%>%filter(month_extracted%in%dormant)%>%dplyr::select(-c(month_extracted))


VARX=list()
VARX$k=1 # indicates that the first two series are modeled
VARX$m=11
VARX$s=2 # sets 2 as the maximal lag order for exogenous series
Model1_growing_78=constructModel(data.matrix(growing_dat_78%>%drop_na),p=4,struct='BasicEN',gran=c(500,10),verbose=FALSE,VARX=VARX,T1=floor(nrow(growing_dat_78)/3),T2=floor(2*nrow(growing_dat_78)/3))
Model1_dormant_78=constructModel(data.matrix(dormant_dat_78%>%drop_na),p=4,struct='BasicEN',gran=c(500,10),verbose=FALSE,VARX=VARX,T1=floor(nrow(dormant_dat_78)/3),T2=floor(2*nrow(dormant_dat_78)/3))
results_growing=cv.BigVAR(Model1_growing_78)
results_dormant=cv.BigVAR(Model1_dormant_78)
# results
# plot(results_growing)
# plot(results_dormant)
# SparsityPlot.BigVAR.results(results_growing)
# SparsityPlot.BigVAR.results(results_dormant)
# 
# predict(results_growing,n.ahead=1, confint=TRUE)
# predict(results_dormant,n.ahead=1, confint=TRUE)
# 
# coef(results_growing)
# coef(results_dormant)

COEF_78=COEF_78%>%bind_rows(coef(results_growing))
COEF_78=COEF_78%>%bind_rows(coef(results_dormant))


```

```{r,fig.width=1.6,fig.height=1.6}
Y1=growing_dat_78%>%drop_na%>%dplyr::select("Depth_cm_b")%>%as.matrix()
X1=growing_dat_78%>%drop_na()%>%dplyr::select(-c("Depth_cm_b"))%>%as.matrix()
# T1=floor(nrow(Y1)/3)
# T2=floor(2*nrow(Y1)/3)
# #X <- matrix(0,nrow=nrow(Y),ncol=ncol(Y))
# BICMSFE <- VARXForecastEval(Y1,X1,p=4,s=2,T1,T2,"BIC",1)##****This part gives the error.
# 
# BICMSFE
# #y=tail(Y1,1198)
# #mon=dat_ws80_2006_2019%>%dplyr::select("month_extracted")
# #mon=mon$month_extracted[((T1+1):T2)]
# y=c(Y1[c(T1+1):T2,1])
# x=c(BICMSFE$pred)
x=c(results_growing@preds)
y=tail(Y1[,1],length(x))
mon=dat_ws77_2006_2019%>%drop_na%>%dplyr::select("month_extracted")
mon=tail(mon$month_extracted,length(x))
preformance_test_depth_78=preformance_test_depth_78%>%bind_cols(data.frame(gof(y, x)))

# plot(x,y,ylim=c(-300,100),xlim=c(-300,100),xlab="predicted",ylab="act",pch=19)
# #title('water table depth y=x, in cm, growing ws78_recent_time',cex.main=.5)
# abline(0,1)
# ggsave(file="water table depth y=x, in cm, growing ws78_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```



```{r,fig.width=1.6,fig.height=1.6}

ggplot( data = data.frame(x,y),aes(x,y)) +
  # stat_poly_line()+
  # stat_poly_eq(use_label(c("eq","adj.R2", "f", "p")),size = 1.5) +
  xlab("predicted")+
  ylab("act")+
  xlim(-300,100)+
  ylim(-300,100)+
 
 # ggtitle("water table depth, in cm, growing ws78")+
  geom_point()+
  theme(plot.title = element_text(size = 8, face = "bold"))
ggsave(file="water table depth, in cm, growing ws78_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```

```{r}

library(ggplot2)

#create residual plot with title and axis labels
ggplot(data = data.frame(x,y), aes(x = x, y = y-x)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE)+
  labs(x='Fitted Values', y='Residuals')+
  theme(plot.title = element_text(size = 14, face = "bold"))
ggsave(file="Residual analysis, water table depth, in cm, growing ws78_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```

```{r,fig.width=1.6,fig.height=1.6}
Y2=dormant_dat_78%>%drop_na%>%dplyr::select("Depth_cm_b")%>%as.matrix()
X2=dormant_dat_78%>%drop_na()%>%dplyr::select(-c("Depth_cm_b"))%>%as.matrix()
# T1=floor(nrow(Y2)/3)
# T2=floor(2*nrow(Y2)/3)
# #X <- matrix(0,nrow=nrow(Y),ncol=ncol(Y))
# BICMSFE <- VARXForecastEval(Y2,X2,p=4,s=2,T1,T2,"BIC",1)##****This part gives the error.
# 
# BICMSFE
# #y=tail(Y2,517)
# #mon=dat_ws80_2006_2019%>%dplyr::select("month_extracted")
# #mon=mon$month_extracted[((T1+1):T2)]
# y=c(Y2[c(T1+1):T2,1])
# x=c(BICMSFE$pred)
x=c(results_dormant@preds)
y=tail(Y2[,1],length(x))
mon=dat_ws77_2006_2019%>%drop_na%>%dplyr::select("month_extracted")
mon=tail(mon$month_extracted,length(x))
preformance_test_depth_78=preformance_test_depth_78%>%bind_cols(data.frame(gof(y, x)))

# plot(x,y,ylim=c(-300,100),xlim=c(-300,100),xlab="predicted",ylab="act",pch=19)
# #title('water table depth y=x, in cm, dormant ws78',cex.main=.5)
# abline(0,1)
# ggsave(file="water table depth y=x, in cm, dormant ws78_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```

```{r,fig.width=1.6,fig.height=1.6}

ggplot( data = data.frame(x,y),aes(x,y)) +
  # stat_poly_line()+
  # stat_poly_eq(use_label(c("eq","adj.R2", "f", "p")),size = 1.6) +
  xlab("predicted")+
  ylab("act")+
  xlim(-300,100)+
  ylim(-300,100)+
  
  #ggtitle("water table depth, in cm, dormant ws78")+
  geom_point()+
  theme(plot.title = element_text(size = 8, face = "bold"))
ggsave(file="water table depth, in cm, dormant ws78.png_recent_time", device = "png",path="pictures_new_well_depth_recent_time")
```

```{r}
library(ggplot2)

#create residual plot with title and axis labels
ggplot(data = data.frame(x,y), aes(x = x, y = y-x)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE)+
  labs( x='Fitted Values', y='Residuals')+
  theme(plot.title = element_text(size = 14, face = "bold"))
ggsave(file="Residual analysis, water table depth, in cm, dormant ws78_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```

```{r,fig.width=1.6,fig.height=1.6}
growing_dat_77=dat_ws77_2006_2019%>%dplyr::select(c(Depth_cm_b,Air_Temp_C,Soil_Temp_,Rainfall,DailyFlow,Solar_rad_,net_rad,PET,RH,Windspeed_,Wind_direc,Vapor_Kpa,month_extracted))%>%filter(month_extracted%in%growing)%>%dplyr::select(-c(month_extracted))%>%drop_na()
dormant_dat_77=dat_ws77_2006_2019%>%dplyr::select(c(Depth_cm_b,Air_Temp_C,Soil_Temp_,Rainfall,Avg_FlowRa,Solar_rad_,net_rad,PET,RH,Windspeed_,Wind_direc,Vapor_Kpa,month_extracted))%>%filter(month_extracted%in%dormant)%>%dplyr::select(-c(month_extracted))%>%drop_na()



VARX=list()
VARX$k=1 # indicates that the first two series are modeled
VARX$m=11
VARX$s=2 # sets 2 as the maximal lag order for exogenous series
Model1_growing_77=constructModel(data.matrix(growing_dat_77%>%drop_na),p=4,struct='BasicEN',gran=c(500,10),verbose=FALSE,VARX=VARX,T1=floor(nrow(growing_dat_77)/3),T2=floor(2*nrow(growing_dat_77)/3))
Model1_dormant_77=constructModel(data.matrix(dormant_dat_77%>%drop_na),p=4,struct='BasicEN',gran=c(500,10),verbose=FALSE,VARX=VARX,T1=floor(nrow(dormant_dat_77)/3),T2=floor(2*nrow(dormant_dat_77)/3))
results_growing=cv.BigVAR(Model1_growing_77)
results_dormant=cv.BigVAR(Model1_dormant_77)
# results
# plot(results_growing)
# plot(results_dormant)
# SparsityPlot.BigVAR.results(results_growing)
# SparsityPlot.BigVAR.results(results_dormant)
# 
# predict(results_growing,n.ahead=1, confint=TRUE)
# predict(results_dormant,n.ahead=1, confint=TRUE)
# 
# coef(results_growing)
# coef(results_dormant)

COEF_77=COEF_77%>%bind_rows(coef(results_growing))
COEF_77=COEF_77%>%bind_rows(coef(results_dormant))
```

```{r,fig.width=1.6,fig.height=1.6}
Y1=growing_dat_77%>%drop_na%>%dplyr::select("Depth_cm_b")%>%as.matrix()
X1=growing_dat_77%>%drop_na()%>%dplyr::select(-c("Depth_cm_b"))%>%as.matrix()
# T1=floor(nrow(Y1)/3)
# T2=floor(2*nrow(Y1)/3)
# #X <- matrix(0,nrow=nrow(Y),ncol=ncol(Y))
# BICMSFE <- VARXForecastEval(Y1,X1,p=4,2,T1,T2,"BIC",1)##****This part gives the error.
# 
# BICMSFE
# #y=tail(Y1,1364)
# #mon=dat_ws80_2006_2019%>%dplyr::select("month_extracted")
# #mon=mon$month_extracted[((T1+1):T2)]
# y=c(Y1[c(T1+1):T2,1])
# x=c(BICMSFE$pred)
x=c(results_growing@preds)
y=tail(Y1[,1],length(x))
mon=dat_ws77_2006_2019%>%drop_na%>%dplyr::select("month_extracted")
mon=tail(mon$month_extracted,length(x))
preformance_test_depth_77=preformance_test_depth_77%>%bind_cols(data.frame(gof(y, x)))

# plot(x,y,ylim=c(-300,100),xlim=c(-300,100),xlab="predicted",ylab="act",pch=19)
# #title('water table depth y=x, in cm, growing ws77',cex.main=.5)
# abline(0,1)
# ggsave(file="water table depth y=x, in cm, growing ws77_recent_time.png", device = "png",path="pictures_new_well_depth")
```



```{r,fig.width=1.6,fig.height=1.6}

ggplot( data = data.frame(x,y),aes(x,y)) +
  # stat_poly_line()+
  # stat_poly_eq(use_label(c("eq","adj.R2", "f", "p")),size = 1.6) +
  xlab("predicted")+
  ylab("act")+
  xlim(-300,100)+
  ylim(-300,100)+
  
  #ggtitle("water table depth, in cm, growing ws77")+
  geom_point()+
  theme(plot.title = element_text(size = 8, face = "bold"))
ggsave(file="water table depth, in cm, growing ws77_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")

```

```{r}
library(ggplot2)

#create residual plot with title and axis labels
ggplot(data = data.frame(x,y), aes(x =x, y = y-x)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE)+
  labs( x='Fitted Values', y='Residuals')+
  theme(plot.title = element_text(size = 14, face = "bold"))
ggsave(file="Residual analysis, water table depth, in cm, growing ws77_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```

```{r,fig.width=1.6,fig.height=1.6}
Y2=dormant_dat_77%>%drop_na%>%dplyr::select("Depth_cm_b")%>%as.matrix()
X2=dormant_dat_77%>%drop_na()%>%dplyr::select(-c("Depth_cm_b"))%>%as.matrix()
# T1=floor(nrow(Y2)/3)
# T2=floor(2*nrow(Y2)/3)
# #X <- matrix(0,nrow=nrow(Y),ncol=ncol(Y))
# BICMSFE <- VARXForecastEval(Y2,X2,p=4,s=2,T1,T2,"BIC",1)##****This part gives the error.
# 
# BICMSFE
# #y=tail(Y2,562)
# #mon=dat_ws80_2006_2019%>%dplyr::select("month_extracted")
# #mon=mon$month_extracted[((T1+1):T2)]
# y=c(Y2[c(T1+1):T2,1])
# x=c(BICMSFE$pred)
x=c(results_dormant@preds)
y=tail(Y2[,1],length(x))
mon=dat_ws77_2006_2019%>%drop_na%>%dplyr::select("month_extracted")
mon=tail(mon$month_extracted,length(x))
preformance_test_depth_77=preformance_test_depth_77%>%bind_cols(data.frame(gof(y, x)))

# plot(x,y,ylim=c(-300,100),xlim=c(-300,100),xlab="predicted",ylab="act",pch=19)
# #title('water table depth y=x, in cm, dormant ws77',cex.main=.5)
# abline(0,1)
# ggsave(file="water table depth y=x, in cm, dormant ws77_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```


```{r,fig.width=1.6,fig.height=1.6}
#par(cex.main=0.2)
ggplot( data = data.frame(x,y),aes(x,y)) +
  # stat_poly_line()+
  # stat_poly_eq(use_label(c("eq","adj.R2", "f", "p")),size = 1.6) +
  xlab("predicted")+
  ylab("act")+
  xlim(-300,100)+
  ylim(-300,100)+
 # ggtitle("water table depth, in cm, dormant ws77")+
  geom_point()+
  theme(plot.title = element_text(size = 8, face = "bold"))
ggsave(file="water table depth, in cm, dormant ws77_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```

```{r}
library(ggplot2)

#create residual plot with title and axis labels
ggplot(data = data.frame(x,y), aes(x = x, y = y-x)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE)+
 labs( x='Fitted Values', y='Residuals')+
  theme(plot.title = element_text(size = 14, face = "bold"))
ggsave(file="Residual analysis, water table depth, in cm, dormant ws77_recent_time.png", device = "png",path="pictures_new_well_depth_recent_time")
```

```{r}
row.names(COEF_77)=c("ws77_daily","ws77_monthly","ws77_growing","ws77_dormant")
row.names(COEF_78)=c("ws78_daily","ws78_monthly","ws78_growing","ws78_dormant")
row.names(COEF_80)=c("ws80_daily","ws80_monthly","ws80_growing","ws80_dormant")
```

```{r}
COEF=cbind(t(COEF_77),t(COEF_78),t(COEF_80))
row.names(COEF)=c("intercept","Y1L1","Y1L2","Y1L3","Y1L4","Air_Temp_C1","Soil_Temp_1","Rainfall1","DailyFlow1","Solar_rad_1","net_rad1","PET1","RH1","Windspeed_1","Wind_direc1","Vapor_Kpa1","Air_Temp_C2","Soil_Temp_2","Rainfall2","DailyFlow2","Solar_rad_2","net_rad2","PET2","RH2","Windspeed_2","Wind_direc2","Vapor_Kpa2")
COEF
```


```{r}
colnames(preformance_test_depth_77)=c("ws77_daily","ws77_monthly","ws77_growing","ws77_dormant")
colnames(preformance_test_depth_78)=c("ws78_daily","ws78_monthly","ws78_growing","ws78_dormant")
colnames(preformance_test_depth_80)=c("ws80_daily","ws80_monthly","ws80_growing","ws80_dormant")
```

```{r}
PERFORMANCE_test_depth=cbind(preformance_test_depth_77,preformance_test_depth_78,preformance_test_depth_80)
PERFORMANCE_test_depth
```
```{r}
as.matrix(PERFORMANCE_test_depth)
```

```{r}
#install.packages("corrplot")
# library(corrplot)
# COEF1=COEF*10000
# COEF2=(COEF1-min(COEF1))/(max(COEF1)-min(COEF1))*100
#corrplot(COEF2, method = 'number') # colorful number
```


```{r}
# df=COEF
# brks <- quantile(df, probs = seq(.05, .95, .05), na.rm = TRUE)
# clrs <- round(seq(255, 40, length.out = length(brks) + 1), 0) %>%
#   {paste0("rgb(255,", ., ",", ., ")")}
# datatable(df) %>% formatStyle(names(df), backgroundColor = styleInterval(brks, clrs))
```






```{r}
# #install.packages('pheatmap') # if not installed already
# library(pheatmap)
# pheatmap(COEF*10000, display_numbers = T,legend_breaks = seq(min(COEF*10000),max(COEF*10000),10))
```


```{r}
Y_77|>pull(Depth_cm_b)|>hist(breaks=200,prob=T)
```

```{r}
Y_78|>pull(Depth_cm_b)|>hist(breaks=200,prob=T)

```
```{r}
Y_80|>pull(Water.Table.Depth.BGS..cm.)|>hist(breaks=200,prob=T)

```


```{r}
length(Y_77|>pull(Depth_cm_b))
length(Y_78|>pull(Depth_cm_b))
length(Y_80|>pull(Water.Table.Depth.BGS..cm.))
```

```{r}
NC_data_1988_2008=read_excel("/Users/alokesh.manna17gmail.com/NSF_internship_2023/raw_data/NC_data/NC_1988_2008.xlsx")
```

```{r}
data.frame(water_table_depth=Y_77|>pull(Depth_cm_b))|>
  mutate(watershed="WS_77")|>
  
bind_rows(data.frame(water_table_depth=Y_78|>pull(Depth_cm_b))|>
  mutate(watershed="WS_78"))|>  
bind_rows( data.frame(water_table_depth=Y_80|>pull(Water.Table.Depth.BGS..cm.))|>
  mutate(watershed="WS_80"))|>
bind_rows( data.frame(water_table_depth=NC_data_1988_2008|>pull(Depth_cm_b))|>
             mutate(water_table_depth=water_table_depth*(-100))|>
  mutate(watershed="NC"))|>
  ggplot(aes(x = water_table_depth)) + 
  geom_histogram(fill = "blue", color = "black") + 
  facet_wrap(~watershed) +
  theme_minimal() +
  labs(title = "Histogram of daily water table depth accross 3  different location in SC and 1 in NC",
       x = "CM",
       y = "Frequency")
```


```{r}
growing_dat_77=dat_ws77_2006_2019%>%dplyr::select(c(Depth_cm_b,Air_Temp_C,Soil_Temp_,Rainfall,Solar_rad_,net_rad,PET,RH,Windspeed_,Wind_direc,Vapor_Kpa,month_extracted))%>%filter(month_extracted%in%growing)%>%dplyr::select(-c(month_extracted))%>%drop_na()
dormant_dat_77=dat_ws77_2006_2019%>%dplyr::select(c(Depth_cm_b,Air_Temp_C,Soil_Temp_,Rainfall,Solar_rad_,net_rad,PET,RH,Windspeed_,Wind_direc,Vapor_Kpa,month_extracted))%>%filter(month_extracted%in%dormant)%>%dplyr::select(-c(month_extracted))%>%drop_na()

growing_dat_78=dat_ws78_2006_2019%>%dplyr::select(c(Depth_cm_b,Air_Temp_C,Soil_Temp,Rainfall_m.x,Solar_rad_,net_rad,PET,RH,Windspeed_,Wind_direc,Vapor_Kpa,month_extracted))%>%filter(month_extracted%in%growing)%>%dplyr::select(-c(month_extracted))
dormant_dat_78=dat_ws78_2006_2019%>%dplyr::select(c(Depth_cm_b,Air_Temp_C,Soil_Temp,Rainfall_m.x,Solar_rad_,net_rad,PET,RH,Windspeed_,Wind_direc,Vapor_Kpa,month_extracted))%>%filter(month_extracted%in%dormant)%>%dplyr::select(-c(month_extracted))

growing_dat_80=dat_ws80_2006_2019%>%dplyr::select(c(Water.Table.Depth.BGS..cm.,Ave_air_T,Soil_Temp_,Rainfall_m,Ave_solar,net_rad,PET,RH,Windspeed_,Wind_direc,Ave_vapor_,month_extracted))%>%filter(month_extracted%in%growing)%>%filter(month_extracted%in%growing)%>%dplyr::select(-c(month_extracted))%>%drop_na()
dormant_dat_80=dat_ws80_2006_2019%>%dplyr::select(c(Water.Table.Depth.BGS..cm.,Ave_air_T,Soil_Temp_,Rainfall_m,Ave_solar,net_rad,PET,RH,Windspeed_,Wind_direc,Ave_vapor_,month_extracted))%>%filter(month_extracted%in%dormant)%>%dplyr::select(-c(month_extracted))%>%drop_na()

```



```{r}
dat_growing_dormant=data.frame(
  water_table_depth=growing_dat_77|>
    pull(Depth_cm_b)
  
)|>mutate(season="growing",
  watershed="WS_77")|>
  
  bind_rows(
    data.frame(
  water_table_depth=dormant_dat_77|>
    pull(Depth_cm_b)
  
)|>mutate(season="dormant",
  watershed="WS_77")
  )|>

bind_rows(
 data.frame(
  water_table_depth=growing_dat_78|>
    pull(Depth_cm_b)
  
)|>mutate(season="growing",
  watershed="WS_78")|>
  
  bind_rows(
    data.frame(
  water_table_depth=dormant_dat_78|>
    pull(Depth_cm_b)
  
)|>mutate(season="dormant",
  watershed="WS_78")
  )
 
  
)|>
  
bind_rows(data.frame(
  water_table_depth=growing_dat_80|>
    pull(Water.Table.Depth.BGS..cm.)
  
)|>mutate(season="growing",
  watershed="WS_80")|>
  
  bind_rows(
    data.frame(
  water_table_depth=dormant_dat_80|>
    pull(Water.Table.Depth.BGS..cm.)
  
)|>mutate(season="dormant",
  watershed="WS_80")
  ))
  
  
```




```{r}
dat_growing_dormant|>
  ggplot(aes(x = water_table_depth)) + 
  geom_histogram(fill = "blue", color = "black") + 
  facet_grid(season~watershed) +
  theme_minimal() +
  labs(title = "Histogram of seasonal variation of water table depth accross 3  different location in SC",
       x = "CM",
       y = "Frequency")
```

##All the 11 variables in beta are written sequence wise and coloured by their importance. starts ##with air temp and end with vapour kpa. 



```{r}
dat77=read.csv("raw_data/dat77_prediction")
dat78=read.csv("raw_data/dat78_prediction")
dat80=read.csv("raw_data/dat80_prediction")
datNC=read.csv("raw_data/datNC_prediction")
```



```{r}
bigdat= dat77|>
  dplyr::select("actual","predicted")|>
  mutate(location="WS_77")|>
  bind_rows(dat78|>
  dplyr::select("actual","predicted")|>
  mutate(location="WS_78"))|>
  bind_rows(dat80|>
  dplyr::select("actual","predicted")|>
  mutate(location="WS_80"))|>
  bind_rows(datNC|>
  dplyr::select("actual","predicted")|>
  mutate(location="WS_NC"))
```


```{r}
library("ggpmisc")
ggplot(bigdat, aes(x = actual, y = predicted)) +
  geom_point(color = "blue") +  # Scatter plot of actual vs predicted
  geom_smooth(method = "lm", se = TRUE, color = "red") +  # Add linear regression line
  stat_poly_eq(formula = y ~ x, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               parse = TRUE) +  # Add regression equation and R^2
  labs(title = "Actual vs. Predicted Values in ws 77, 78, 80 in SC and a ws in NC",
       x = "Actual Values",
       y = "Predicted Values") +
  facet_wrap(~ location) +
  theme_minimal() 
```

